---
title: "Export"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo =FALSE, message=FALSE}
library(shortestpath)
library(knitr)
library(igraph)
set.seed(22)

graph <- make_ring(4) %>%
    set_edge_attr("weight", value = c(2,2,3,1) ) %>%
    as.spgraph()

```
The shortest path package offers different functions to visualize the optimization steps of each algorithm. Moreover, it also provides features to export the results by creating LaTeX code (tables and tikz plots) out of spgraphs.

## plot function
By simply plotting an object of class spgraph with the normal `plot()` function, the `plot.spgraph()` will be called automatically. The `plot.spgraph()` is a wrapper around `plot.igraph()` providing convenience features for shortest path graph plotting. Amongst other things it allows the user to set characteristics and corresponding values for the vertices and eges within the graph.

```{r,fig.width=4, fig.height=3, comment = NA}
d <- dijkstra(graph, "A", "C")

plot(d$last, 
     vertex.color.by = "type", vertex.color = rainbow(3),
     vertex.frame.color.by = "set", vertex.frame.color = rainbow(3),
     edge.color.by = "manual", edge.color = c("darkgreen","pink","black","blue")
     )
```  
  
In order to set the attributes for the vertices and edges we first have to define based on which characteristics the different values should be assigned to the corresponding graph elements. It is important to know that each graph element can have different characteristics.   
In this example the vertex characteristics can be seperated into "set","type" and "manual". The characteristic "set" is based on the vertex attribute "set", which consist of the values "known", "unknown" and "front".   
Whereas for the characteristics of the edges we can only choose "shortest path" or "manual".  
The next parameter defines which colors should be used.
  
Further explanation and an overview of the different parameters can be found in the R documentation `plot.spgraph` of the shortestpath package.

## toLatexTable
The toLatexTable function creates LaTeX code consisting of the minimum distances and shortest path predecessors table for each step of the Dijkstra algorithm  

```{r,fig.width=7, fig.height=6, comment = NA}
par(mfrow = c(2, 2))
plot(d)
toLatexTable(d)
```

## toLatexGraph
The toLatexGraph functions creates LaTeX code (tikz plot) consisting of the spgraph.  
```{r,fig.width=7, fig.height=6,comment = NA}
toLatexGraph(d)
```
