---
title: "generating graphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{knitr,shortestpath}
---
```{r, echo =FALSE, message=FALSE}
library(shortestpath)
library(knitr)
library(igraph)
set.seed(30)
```
There are more than one way to create a proper graph in r.

## igraphs's make_graph
The igraph package provides a function "make_graph", which creates an igraph graph from a list of edges, or notable graph.

```{r,fig.width=4, fig.height=3}
par(mar=c(0,0,0,0))
g <- make_graph("House")
plot(g)
```
In this example the notable graph "House" is used. A list of notable graphs can be found in the R documentation of igraphh.  

## igraphs's graph_from_adjacency_matrix
Another function of igraph - the graph_from_adjancency_matrix - creates a graph out of an adjancency matrix. 

```{r,fig.width=4, fig.height=3}
par(mar=c(0,0,0,0))
square <- matrix(c(
    0,1,0,1,
    1,0,1,0,
    0,1,0,1,
    1,0,1,0
    ), nrow=4) %>%
graph_from_adjacency_matrix(mode="undirected", weighted=TRUE)
V(square)$name <- LETTERS[1:4]
V(square)$x <- c(0, 1, 1, 0)
V(square)$y <- c(0, 0, 1, 1)
plot(square)
```

## shortestpath's randomGraph
This method generates a random graph only by specyfing the number of nodes and the connectivity (k). The result is a not too  graph which is appropriate for exercises

```{r,fig.width=4, fig.height=3}
par(mar=c(0,0,0,0))
g = randomGraph(no.of.nodes = 6, k = 2.5, euclidean = FALSE)
plot(g)
```
The result of the function looking graphs, which are not too generated and therefore suitable for exercises

To get "nice" random graphs, we create a sample_k_regular graph
    # with a degree of k*2 and then randomly remove edges. This creates graphs
    # which aren't too degenerated.


