---
title: "Getting Started with shortestpath"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{knitr,shortestpath}
---
```{r, echo =FALSE, message=FALSE}
library(shortestpath)
library(igraph)
library(knitr)
set.seed(22)
```

##Intro
The shortestpath package provides functions to solve the shortest path problem by using the Dijkstra-, Floyd-Warshall, A*Search and Bellman-Ford algorithm. Additionally, the package offers 
the tabular and graphical visualization of the shortest path optimization steps. 

## Create a Graph
As a first step a graph need to be created. There are many ways to create graph, which are explained in detail in generating-graphs vignette. In this example the igraph function `make_ring()` is used in order to create a undirected graph `g` with positive edge weights. 
```{r}
g <- make_ring(4) %>%
    set_edge_attr("weight", value = c(2,2,3,1) ) %>%
    as.spgraph()
```
The parameter "4" indicates the number of vertices of the generated graph.

## Dijkstra algorithm 
Use the Dijkstra algrithm to find the shortest path between the two vertices "A" and "C" in the graph `g`.
```{r}
d <- dijkstra(g, "A", "C")
```
The variable `d` contains a list of igraphs (called spresults). These igraphs have additional attributes and are therefore called shortest path graphs (spgraphs). Each spgraph represents one optimization step of the Dijkstra algorithm. Details of the data structure can be found in the data-structure vignette.

## Plot the graph 
Use the normal `plot()` function to print all the steps of the Dijkstra algorithm. 
```{r,fig.width=7, fig.height=6}
par(mfrow = c(2, 2))
plot(d)
```
A detailed explanation of the single optimization steps and the `plot()` can be read in the processing and export vignette.  

<br> 
Additionally, the minimum distance table can be printed by calling the attribute `min_dists` of each step (spgraph).
```{r}
 for(step in d){ 
   print(step$min_dists)
   cat("\n")
   }
```
