---
title: "01 Getting Started with shortestpath"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{01 Getting Started with shortestpath}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{knitr,shortestpath}
---
```{r, echo = FALSE, message = FALSE}
library(shortestpath)
library(knitr)
opts_chunk$set(dev = "svg")
```
```{r, echo = FALSE}
set.seed(22)
```

## Intro
The shortestpath package provides a set of functions to execute and visualize algorithms that solve shortest path problems. For example, shortestpath can plot individual steps of Dijkstra or A*-Search; print distance tables for each step of the Floyd-Warshall or Bellman-Ford algorithm or export graphs to human-readable TikZ.

Before we start, there are a two important bits to know. First, shortestpath heavily integrates into the existing R ecosystem: Our graph data structures are based on (and interoperable with) the excellent [igraph](http://igraph.org/) library, so if you know igraph, you'll notice lots of familiarities. We also make heavy use of [magrittr](https://github.com/smbache/magrittr)'s `%>%` pipes to structure our examples - please check out their very short [introduction](https://github.com/smbache/magrittr) first if you never used magrittr before.

## The Big Picture

Your shortestpath workflow will generally look as follows:

1. Graph Import/Generation
2. Running a Shortest Path Algorithm
3. Results Visualization/Export

We will walk through each of these steps here - for more details, each step has a dedicated vignette!

## Create a Graph
Before we start finding shortest paths, we need a graph! There are many ways to create graphs which are explained in detail in the 'Generating Graphs' vignette. For a quick example, we use the igraph function [`make_ring()`](http://igraph.org/r/doc/make_ring.html) to create an undirected ring graph `g` with positive edge weights. Next, `as.spgraph()` converts the graph into a shortest path graph. An spgraph is still a valid igraph, but it is also guaranteed to have a limited set of additional attributes (e.g. fixed vertex positions, so that plots render consistently over optimization steps).

```{r, message=FALSE}
library(igraph)

g <- make_ring(4) %>%
    set_edge_attr("weight", value = c(4,1,3,3) ) %>%
    as.spgraph()

plot(g)
```

## Find the shortest path using Dijkstra's algorithm
We now want to find the the shortest path from A to C using Dijkstra's algorithm.
```{r}
d <- dijkstra(g, "A", "C")
print(d)
```

Dijsktra (and all other algorithms) returns an `spresults` object. In a nutshell (we will discuss them in more detail in the 'Data Structures' vignette), an ``spresults`` object is a list of ``spgraph``s, with each entry representing a single step of your algorithm.

## Plot the graph 
Use the normal `plot()` function to print all the steps of the Dijkstra algorithm. 
```{r}
par(mfrow = c(2, 2))
plot(d)
```

The plot command can be configured with a wide range of settings - we will discuss this in the vignette 'Exporting Results'!

Of course, you can also work with the returned datastructure manually: In the following example, we obtain the minimum distance table after each step of Dijkstra.
```{r}
lapply(d, function(step) {
    step$min_dists
})
```
