---
title: "Generating graphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating Graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{knitr,shortestpath}
---
```{r, echo = FALSE, message = FALSE}
library(shortestpath)
library(knitr)
opts_chunk$set(dev = "svg")
```
```{r, echo = FALSE}
set.seed(30)
```

To leverage the existing R ecosystem for graphs, shortestpath is closely coupled with igraph and accepts any valid igraph object as input. Additionally, we provide a configurable random graph generation function that is particularly suited for exercises.

In this vignette, we will first walk through the three most common ways to generate graphs and then show how to remove intersecting edges for visually appealing graphs.

## igraphs's make_graph
The igraph package provides a function `make_graph()`, which can be used to create an igraph from a list of edges or a string specifying a well-defined notable graph. Calling `as.spgraph()` casts the igraph to an spgraph. This adds alphabetical vertex names, uniform edge weights and fixed vertex coordinates by default.

```{r, message=FALSE}
library(igraph)
g <- make_graph("House") %>%
    as.spgraph()

plot(g)
```
  
A list of predefined graphs can be found in the documentation of `make_graph()`.  

## igraphs's graph_from_adjacency_matrix
igraph's `graph_from_adjancency_matrix()` creates a graph from an adjancency matrix. 

```{r}
square <- matrix(c(
    0,1,0,1,
    1,0,1,0,
    0,1,0,1,
    1,0,1,0
    ), nrow=4) %>%
    graph_from_adjacency_matrix(mode="undirected") %>%
    as.spgraph()
plot(square)
```

## shortestpath's randomGraph
This method generates a random graph only by specifying the number of nodes and the average connectivity. The result is a graph that is neither too uniform nor too degenerated and suitable for exercises.

```{r}
g <- randomGraph(no.of.nodes = 6, k = 2.5, euclidean = FALSE) 
plot(g)
```
  
## Remove overlapping edges
By creating a new graph with the above mentioned functions it can happen that some vertices are placed in such a position in the layout that their edges intersect. In order to create visually appealing graphs, shortestpath's `removeIntersectingEdges()` can be used.  
The function tries to eliminate edge intersections in a graph by replacing overlapping edges with new -- non-overlapping -- ones. Newly added edges will not have any edge attributes.

```{r, echo = FALSE}
set.seed(1)
```
```{r}
g <- make_graph("Cubical") %>%
    as.spgraph()
plot(g)

g <- g %>%
    removeIntersectingEdges(relayout = FALSE) %>%
    setUniformEdgeWeights()
plot(g)
```
  
If the `relayout` parameter is `TRUE`, the Fruchterman-Reingold layout algorithm is re-applied after removing edges. This may introduce new intersections, but generally improves the quality of the graph.
  
It is important to know that this is a fuzzy algorithm, there is no guarantee that all intersections are (or can be) removed. If relayout is set to `FALSE`, however, it is guaranteed that a warning will be emitted if not all intersections could be removed.
