---
title: "Exporting Results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exporting Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE, message = FALSE}
library(shortestpath)
library(knitr)
opts_chunk$set(dev = "svg")
```
```{r, echo = FALSE}
set.seed(22)
```
The shortest path package offers different functions to visualize the optimization steps of each algorithm. Moreover, it also provides features to export the results to LaTeX tables and TikZ plots.

## plot function
spgraph and spresults objects can be plotted using R's builtin `plot()` function (which internally calls `plot.spgraph()`). `plot.spgraph()` is a wrapper around `plot.igraph()`, providing additional features for shortest path graph plotting and a generally more visually appealing output. Amongst other things, it allows the user to set vertex/edge visuals depending on the underlying graph characteristics:

- `vertex.color.by` denotes which attribute should be used to set the visual
- `vertex.color` specifies the color palette, or, if `.by` is `"manual"`, a color for each vertex.

For example, setting `vertex.color.by="set"` for an A*-Search graph colors each vertex by its current set (processed/at the current front/unknown).

```{r, message=FALSE}
library(igraph)
graph <- make_ring(4) %>%
    as.spgraph() %>%
    setRandomEdgeWeights()

d <- dijkstra(graph, "A", "C")

plot(d$last, 
     vertex.color.by = "type", 
     vertex.color = heat.colors(3),
     vertex.frame.color.by = "manual", 
     vertex.frame.color = c("darkgreen","pink","black","blue"),
     edge.color.by = "shortestpath", 
     edge.color = c("gray","green")
     )
```  
  
An overview of the different parameters can be found in the R help for `plot.spgraph`.

### Animations

It is also very useful to create animated GIFs, videos or LaTeX plots using the R `animation` package. In the following example, we use it in combination with `Cairo` to obtain a high-quality antialiased video of A*-Search.

```{r, eval=FALSE}
library(Cairo)
library(animation)

grid_search <- make_lattice(length = 7, dim = 2) %>%
    delete_edges(c(seq(21,25,2),seq(37,63,13),seq(60,64,2))) %>%
    as.spgraph() %>%
    setVertexCoordinatesFromLayout(
        layout=on_grid(width=7, height=7)
    ) %>%
    aStarSearch("22", "28", 
        distance.heuristic = function(...) euclidean.vertex.distance(...) ** 2
    )

saveVideo({
    CairoPNG(filename = ani.options('img.fmt'))
    plot(grid_search, vertex.label="", edge.label="")
    dev.off()
}, video.name="a-star.mp4", use.dev=FALSE, ani.type="png")
```

## toLatexTable
The toLatexTable function creates LaTeX code consisting of the minimum distances and shortest path predecessors table for each step of the Dijkstra algorithm  

```{r, comment = NA}
toLatexTable(d)
```

## toLatexGraph
The toLatexGraph functions creates a human-readable TikZ plot of the spgraph. The emphasis of this function is to provide (in contrast to R's TikZ plot device) readable TikZ code. 
```{r, comment = NA}
toLatexGraph(d)
```
