---
title: "Data Structures"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{05 Data Structures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{knitr,shortestpath}
---
```{r, echo = FALSE, message = FALSE}
library(shortestpath)
library(knitr)
opts_chunk$set(dev = "svg")
```
```{r, echo = FALSE}
set.seed(22)
```
The shortestpath package uses igraph's graph objects to represent graphs. However, additional attributes are added to the igraph graph in order to show the steps of finding a shortest path by the corresponding algorithm. The resulting object is called spgraph. In other words, spgraph is a subclass of igraph with required additional attributes.

## spgraph
The spgraph is an igraph graph with the following additional attributes:
```{r, message=FALSE}
library(igraph)
graph <- make_graph("House") %>%
    as.spgraph()

print(graph)
```

## spresults

Each algorithm (in this case dijkstra) returns an spresults object. An spresults object is a list of spgraphs, with each graph depicting a single iteration of the algorithm.

```{r}
results <- dijkstra(graph, "A", "D")
print(results)
str(results, list.len=2)
head(lapply(results, function(step) step$min_dists), 3)
```

As a convenience, you can access graph attributes of the last iteration using `spresults$attr`:
```{r, results = "hide"}
# These are equivalent:
results[[length(results)]]$min_dists == results$min_dists
```

Probably the most interesting attributes are `min_dists` and `shortest_path_predecessors`:

```{r}
results$min_dists
results$shortest_path_predecessors
```
The shortest path predecessors matrix consists of `igraph.vs` lists, which allows to reconstruct the actual shortest path. Each algorithm is able to indentfy *all* shortest paths: In the example above, there are two shortest paths from A to D.  

To access individual elements of a matrix of lists in R, [you have to use double brackets](http://stackoverflow.com/questions/30007890/how-to-create-a-matrix-of-lists-in-r):
```{r}
results$shortest_path_predecessors[["A","D"]]
```

### Getting all shortest paths

You can use shortestpath's `getShortestPaths()` method to get a list of all shortest paths:

```{r}
getShortestPaths(results)
```

## Graph Modification
The shortestpath package provides many graph modification functions for spgraphs and igraphs that are used internally. Further details can be found in the R documentation: `?graphModification`

```{r}
par(mar = c(0,0,0,0))
g <- random.graph.game(10, 0.5) %>%
  setAlphabeticalVertexNames %>%
  setRandomEdgeWeights()

plot(g)
E(g)$weight
```
  
